---
  output: 
    blastula::blastula_email
  title: "Crowd Forecast Updates"
  fig_caption: false
---

```{r echo=FALSE, message=FALSE, warning=FALSE}

knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE, 
                      warning = FALSE)


library(here)
library(scoringutils)
library(dplyr)
library(data.table)
library(covid.german.forecasts)
library(kableExtra)
```


```{r load-data, echo=FALSE, message = FALSE, warning = FALSE}
# load all data ----------------------------------------------------------------
root_dir <- here::here("crowd-forecast", "processed-forecast-data")
file_paths_forecast <- here::here(root_dir, list.files(root_dir))

prediction_data <- purrr::map_dfr(file_paths_forecast, 
                                  .f = function(x) {
                                    data <- data.table::fread(x) %>%
                                      dplyr::mutate(target_end_date = as.Date(target_end_date), 
                                                    submission_date = as.Date(submission_date), 
                                                    forecast_date = as.Date(forecast_date))
                                  }) %>%
  dplyr::mutate(target_type = ifelse(grepl("death", target), "death", "case")) %>%
  dplyr::rename(prediction = value) %>%
  dplyr::mutate(forecast_date = as.Date(submission_date)) %>%
  dplyr::rename(model = board_name) %>%
  dplyr::filter(type == "quantile", 
                location_name %in% c("Germany", "Poland")) %>%
  dplyr::select(location, location_name, forecast_date, quantile, prediction, model, target_end_date, horizon, target, target_type)

# filter forecasters such that only those are kept that have submitted forecasts
# in the last four weeks
keep_fc <- prediction_data %>%
  dplyr::filter(forecast_date > (Sys.Date() - 4 * 7)) %>%
  dplyr::pull(model) %>%
  unique()

prediction_data <- dplyr::filter(prediction_data, 
                                 model %in% keep_fc)

files <- list.files(here::here("data-raw"))
file_paths <- here::here("data-raw", files[grepl("weekly-incident", files)])
names(file_paths) <- c("case", "death")

truth_data <- purrr::map_dfr(file_paths, readr::read_csv, .id = "target_type") %>%
  dplyr::rename(true_value = value) %>%
  dplyr::mutate(target_end_date = as.Date(target_end_date)) %>%
  dplyr::arrange(location, target_type, target_end_date) %>%
  dplyr::filter(location_name %in% c("Germany", "Poland"))


data <- scoringutils::merge_pred_and_obs(prediction_data, truth_data, 
                                         join = "full")

```


### Dear `r name`,

Good to have you as a forecaster! Here is this weeks reminder and an update about what has happened lately. You're receiving this message because you gave us your email address when you registered for the [Epiforecasts Crowd Forecast App](https://cmmid-lshtm.shinyapps.io/crowd-forecast/). 

If you want to lean more about this project, have a look at our [github repo](https://github.com/epiforecasts/covid.german.forecasts). If you want to learn more about the German and Polish Forecast Hub, where we submit the forecasts to, have a look [here](https://kitmetricslab.github.io/forecasthub/forecast). 

<br>

### The European Forecast Hub

On March 8th, the [European Forecast Hub](https://covid19forecasthub.eu/) will launch. Our crowd forecast app will therefore be expanded to include a set of 32 countries. If you like you can stick to forecasting Germany and Poland, but more observations will be available. We'll send more detailed information around next week. 

<br>

### Introduction video

To help new forecasters get used to the user interface, we created a [short video](https://www.youtube.com/watch?v=NzZkNxXFgm8) that explains how to make a forecast. If you like, check it out. 

<br>

### Reminder for the weekly forecasts

We **update our data every Saturday at 6.30pm CET**. You can then make a forecast until **2pm CET (1pm UK time) on Tuesday** using our [forecast app](https://cmmid-lshtm.shinyapps.io/crowd-forecast/). We would be very happy to have a forecast from you this week!


<br>

### Forecast and Performance overview

But now, let's have a look at last weeks forecasts and performance!

In the following we put together a short overview for you. You can find more detailed information on our performance board [here](https://epiforecasts.io/covid.german.forecasts). 

*Here are the latest EpiExpert ensemble forecasts and next to them your own forecasts
if you have submitted any.*

```{r plot-forecasts, echo=FALSE, out.width="49%", fig.show="hold"}

if (weekdays(Sys.Date()) == "Monday") {
  forecast_date <- covid.german.forecasts::latest_weekday(Sys.Date() - 7)
} else {
  forecast_date <- covid.german.forecasts::latest_weekday(Sys.Date())
}


filter_truth <- list('target_end_date >= Sys.Date() - 14 * 7')
filter_forecasts <- list("model %in% 'EpiExpert-ensemble'", 
                         paste0("forecast_date == '", forecast_date, "'"))

plot_predictions(data = data, 
                 x = "target_end_date",
                 filter_truth = filter_truth, 
                 filter_forecasts = filter_forecasts,
                 facet_formula = target_type ~ location_name, 
                 facet_wrap_or_grid = "grid") + 
  ggplot2::labs(x = "Date", title = "Ensemble Forecasts") + 
  ggplot2::theme(legend.position = "bottom")

filter_truth <- list('target_end_date >= Sys.Date() - 14 * 7')
filter_forecasts <- list(paste0("model %in% '", board_name, "'"), 
                         paste0("forecast_date == '", forecast_date, "'"))

p <- plot_predictions(data = data, 
                 x = "target_end_date",
                 filter_truth = filter_truth, 
                 filter_forecasts = filter_forecasts,
                 facet_formula = target_type ~ location_name, 
                 facet_wrap_or_grid = "grid") 
if (!is.null(p)) {
  p <- p + 
  ggplot2::labs(x = "Date", title = "Your Forecasts") + 
  ggplot2::theme(legend.position = "bottom")
  print(p)
}


```

<br>

*Here is the list with the current top 5 forecasters (based on the last 5 weeks): *

```{r rank-table, echo=FALSE, fig.align="center"}
cut_off <- covid.german.forecasts::latest_weekday(Sys.Date()) - 5 * 7
df <- data %>%
  filter(forecast_date >= cut_off)
  
table <- eval_forecasts(data, summarise_by = c("model"), 
                        compute_relative_skill = TRUE)

table %>%
    arrange(relative_skill) %>%
    dplyr::select(-scaled_rel_skill) %>%
    mutate_if(is.numeric, round, 2) %>%
    dplyr::rename(wis = interval_score,
                  underpred = underprediction,
                  overpred = overprediction,
                  cvrage_dev = coverage_deviation,
                  rel_skill = relative_skill) %>%
  select(model, rel_skill) %>%
  head(5) %>%
  knitr::kable(format = "html") %>% 
  kable_styling(bootstrap_options = "striped", full_width = F)



```

<br>

*And here is a plot with the rank by date and target of all active forecasters over time:*

```{r rank-over-time, echo=FALSE, fig.align = "center", fig.width=10, fig.height=8, out.width="80%"}
scores <- scoringutils::eval_forecasts(data,
                                         summarise_by = c("model", 
                                                          "location_name", 
                                                          "target_type",
                                                          "forecast_date"), 
                                       compute_relative_skill = FALSE)

ranked_scores <- scores %>%
  dplyr::mutate(forecast_date = as.Date(forecast_date)) %>%
  dplyr::group_by(forecast_date, location_name, target_type) %>%
  dplyr::mutate(num_forecasts = dplyr::n(), 
                rank = rank(interval_score, ties.method = "average",
                                           na.last = NA), 
                standard_rank = round((1 - (rank - 1) / (num_forecasts - 1)) * 100)) %>%
  dplyr::ungroup()

plot <- score_heatmap(ranked_scores, metric = "rank", 
                      x = "forecast_date",
                      facet_wrap_or_grid = "grid",
                      facet_formula = location_name ~ target_type) + 
  ggplot2::theme(legend.position = "bottom") +
    ggplot2::scale_fill_gradient(low = "white", high = "coral")

print(plot)
```

<br>

If you have any further questions or feedback, please don't hesitate to reach out. You can drop us an email or submit an issue to our [github repo](https://github.com/epiforecasts/covid.german.forecasts)

Thank you very much for your support and participation. We very much appreciate it!

Best wishes, 

Nikos for the Epiforecasts team

<br>

PS: If you don't want to receive future emails simply write me a quick message. We're working on a way that allows you to unsubscribe from within the app, but that is work in progress. Baby steps, but we're getting there. 

PPS: We solved our quota issues. Thank you very much for your helpful comments last week!
