---
title: "Analysis paper"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(covid.german.forecasts)
library(here)
library(magrittr)
library(scoringutils)
library(knitr)
library(kableExtra)
library(data.table)
library(ggplot2)

# load and set up data
cases <- combined_data[target_type == "case" & 
                         as.Date(forecast_date) %in% forecast_dates$cases]
deaths <- combined_data[target_type == "death" & 
                         as.Date(forecast_date) %in% forecast_dates$deaths]
data <- rbindlist(list(cases, deaths))
```

## Figure with 2 week ahead forecasts and scores

-[x] create plots
-[x] combine plots into one
-[ ] shade dates that are excluded in scores
-[ ] maybe add density plot on the y axis
-[ ] maybe use mean scores ratio instead of WIS for plot with score over time

### Forecasts vs. truth data
```{r}
df <- combined_data[as.Date(forecast_date) %in% forecast_dates$unfiltered &
                      grepl("2", target) & 
                      quantile %in% c(0.025, 0.25, 0.75, 0.975)]
             
df <- dcast(df, ... ~ quantile, value.var = "prediction")

df %>%
  ggplot(aes(y = true_value, x = target_end_date)) + 
  geom_linerange(aes(ymin = `0.025`, ymax = `0.975`, color = model), 
                 size = 1.5,
                 alpha = 0.4,
                 position = position_dodge(width = 3)) + 
  geom_linerange(aes(ymin = `0.25`, ymax = `0.75`, color = model), 
                 size = 1.5,
                 alpha = 0.8,
                 position = position_dodge(width = 3)) + 
  labs(y = "value", x = "date") + 
  geom_point() + 
  geom_line() + 
  theme_light() + 
  facet_grid(target_type ~ location_name, scales = "free") + 
  theme(legend.position = "bottom")

ggsave(here("analysis", "plots", "truth-and-2-wk-ahead.png"))

```


### plot with interval score

```{r}
# plot with scores over time ---------------------------------------------------
scores_target_forecastdate <- eval_forecasts(
  data = data[grepl("2", target)],
  summarise_by = c("model", "location_name", "target_type", "forecast_date"), 
  compute_relative_skill = TRUE, 
  baseline = "Baseline"
)

scores_target_forecastdate  %>%
  ggplot(aes(x = as.Date(forecast_date), y = interval_score, colour = model)) + 
  geom_line() + 
  geom_point() + 
  theme_light() + 
  facet_grid(target_type ~ location_name, scales = "free") +
  theme(legend.position = "bottom") +
  labs(y = "weighted interval score", x = "date") 

ggsave(here("analysis", "plots", "wis-over-time.png"))
```

### same plot with mean score ratio
not sure why there is a whole in the plot...
```{r}
scores_target_forecastdate  %>%
  ggplot(aes(x = as.Date(forecast_date), y = scaled_rel_skill, colour = model)) + 
  geom_line() + 
  geom_point() + 
  theme_light() + 
  facet_grid(target_type ~ location_name, scales = "free") +
  theme(legend.position = "bottom") +
  labs(y = "mean score ratio", x = "date") 
```

---

## Summary table with scores
-[] create summary table for scores visualised before
-[] exclude Christmas period
-[] use whole period for cases
-[] use only period for which we have all models for deaths
-[] also add 50% and 90% coverage in table


```{r}
# create helper function to make table
make_score_table <- function(scores) {
  scores <- as.data.table(scores)
  setnames(scores, 
           old = c("target_type", "interval_score", "aem", "relative_skill", "scaled_rel_skill"), 
           new = c("target", "WIS", "absolute error", "rel. skill", "scaled rel. skill"))
  
  table <- scores[, .SD,  .SDcols = !c("coverage_deviation", "rel. skill")][
    order(target, `scaled rel. skill`)
  ] %>%
    kable(format = "markdown")
  return(table)
}
```


### 2 weeks ahead
```{r }
# scores by target type for 2 week ahead forecasts
eval_forecasts(
  data[grepl("2", target)], 
  summarise_by = c("model", "target_type"), 
  compute_relative_skill = TRUE, 
  baseline = "Baseline"
) %>%
  make_score_table()
```

### 4 weeks ahead
```{r }
# scores by target type for 4 week ahead forecasts
eval_forecasts(
  data[grepl("4", target)], 
  summarise_by = c("model", "target_type"), 
  compute_relative_skill = TRUE, 
  baseline = "Baseline"
) %>%
  make_score_table()

```


---

## Look more closely at relative skill

---

## Analyse distribution of model rankings
- one possibility would be to take the ridge plot from Cramer et al from the Hub

---

## Look at WIS contributions
```{r}
scores_target <- eval_forecasts(
  data[grepl("2", target)], 
  summarise_by = c("model", "target_type"), 
  compute_relative_skill = TRUE
) 

wis_components(scores_target, 
               facet_formula = ~ target_type, 
               scales = "free_x", 
               x_text_angle = 0) + 
  theme(legend.position = "bottom") + 
  coord_flip() + 
  labs(title = "Components of the weighted interval score for 2-week ahead forecasts", 
       x = NULL)

ggsave(here("analysis", "plots", "wis-components.png"))
```

---

## Look more closely at interventions and change points

---

## Look at the effect of data exclusions
- look at first period, second period etc
```{r, eval = FALSE}
scores <- list()
# scores for restricted data set
scores_full <- eval_forecasts(data, 
                              summarise_by = c("model", "target_type", "target"), 
                              compute_relative_skill = TRUE, 
                              baseline = "Baseline")

scores[["full"]] <- 
  scores_full[, .(model, target_type, scaled_rel_skill, interval_score, target,
                  scenario = "full")]

res <- rbindlist(scores)
res %>%
  ggplot(aes(x = scenario, y = scaled_rel_skill, color = model, group = model)) + 
  geom_point() + 
  facet_wrap( ~ target, scales = "free", ncol = 2) + 
  theme(legend.position = "bottom")
```

---

## Look more closely at calibration and sharpness